### 2018-04-25 contents/abstract.tex
Updates from ShareLaTeX

\chapter{Abstract}

Idiopathic (Classic) Trigeminal Neuralgia (TN) is a facial neuropathic pain syndrome characterized by paroxysmal, shock-like pain condition affecting one or more of the three trigeminal nerve (CNV) branches. TN is believed to be associated with nerve-vascular compression in the CNV root-entry-zone, but its pathophysiology is still unclear. Single-tensor diffusion tensor neuroimaging (DTI) studies of CN V revealed diffusivity changes in the cistern segment. However, the portion of the nerve within the brainstem remained elusive due to the limits of DTI, and the lack of reliable software to researchers. Thereby the study of TN using diffusion imaging is time-consuming and error-prone due to the complete manual involvement in all steps of data process, placement of region-of-interest (ROI), and tractography parameter tuning, and data analysis. 
There are two aims of this thesis: 1) to develop a fully-automated software framework to collate and analyze diffusion tractography to reduce human error and increase iteration speed of experiments. 2) to apply the method developed to the analysis of the trigeminal system in a more substantial patient group, and apply state-of-the-art diffusion imaging methods to advance the study of TN further. 
The specific aims are: a) establish the feasibility of high angular resolution diffusion tractography of brainstem CNV; b) establish the best way to minimize DWI to T1 co-registration error allow non-linear registration of tractography across multiple subjects; c) Create the software framework to generate and quantify tractography at the group level; d) apply the methodology to the reconstruction and quantification of the trigeminal sensory pathway. 
Towards these goals, in Study I, we establish the feasibility of applying spherical deconvolution tractography to delineate the full course of CNV from the trigeminal nucleus and demonstrate that TN is uniquely identified by disruptions in the cistern/REZ, while MS-TN can be distinguished by disruptions in the brainstem course of the nerve. In Study II, we determine that the best T1-DWI co-registration scalar is the Mean DWI image, and that the registration can potentially be improved with a novel spherical harmonic parameter scalar imaged called Anisotropic Power image. In Study III, we present the Selective Automated Group Integrated Tractography (SAGIT) software framework. Finally, in Study IV, we deploy end-to-end machine-learning TN classification to auto-discovers diffusivity disruptions in the cistern/REZ CNV, the trigeminopontothalamic decussaion, and thalamocortico S1 pathway. 
In sum, this thesis presents a detailed road-map of both the development and application of a novel end-to-end diffusion tractography machine learning classification. The application of this system to TN revealed specific diffusivity changes in all three segments of the trigeminal CN V, pontine decussation, and S1  white matter pathways, and pin-points the exact location of the diffusivity disruption at millimetre level.

>>>>

\begin{abstract}
\addcontentsline{toc}{chapter}{Abstract}
Idiopathic (Classic) Trigeminal Neuralgia (TN) is a facial neuropathic pain syndrome characterized by paroxysmal, shock-like pain condition affecting one or more of the three trigeminal nerve (CNV) branches. TN is believed to be associated with nerve-vascular compression in the CNV root-entry-zone, but its pathophysiology is still unclear. Single-tensor diffusion tensor neuroimaging (DTI) studies of CN V revealed diffusivity changes in the cistern segment. However, the portion of the nerve within the brainstem remained elusive due to the limits of DTI, and the lack of reliable software to researchers. Thereby the study of TN using diffusion imaging is time-consuming and error-prone due to the complete manual involvement in all steps of data process, placement of region-of-interest (ROI), and tractography parameter tuning, and data analysis. 
There are two aims of this thesis: 1) to develop a fully-automated software framework to collate and analyze diffusion tractography to reduce human error and increase iteration speed of experiments. 2) to apply the method developed to the analysis of the trigeminal system in a more substantial patient group, and apply state-of-the-art diffusion imaging methods to advance the study of TN further. 
The specific aims are: a) establish the feasibility of high angular resolution diffusion tractography of brainstem CNV; b) establish the best way to minimize DWI to T1 co-registration error allow non-linear registration of tractography across multiple subjects; c) Create the software framework to generate and quantify tractography at the group level; d) apply the methodology to the reconstruction and quantification of the trigeminal sensory pathway. 
Towards these goals, in Study I, we establish the feasibility of applying spherical deconvolution tractography to delineate the full course of CNV from the trigeminal nucleus and demonstrate that TN is uniquely identified by disruptions in the cistern/REZ, while MS-TN can be distinguished by disruptions in the brainstem course of the nerve. In Study II, we determine that the best T1-DWI co-registration scalar is the Mean DWI image, and that the registration can potentially be improved with a novel spherical harmonic parameter scalar imaged called Anisotropic Power image. In Study III, we present the Selective Automated Group Integrated Tractography (SAGIT) software framework. Finally, in Study IV, we deploy end-to-end machine-learning TN classification to auto-discovers diffusivity disruptions in the cistern/REZ CNV, the trigeminopontothalamic decussaion, and thalamocortico S1 pathway. 
In sum, this thesis presents a detailed road-map of both the development and application of a novel end-to-end diffusion tractography machine learning classification. The application of this system to TN revealed specific diffusivity changes in all three segments of the trigeminal CN V, pontine decussation, and S1  white matter pathways, and pin-points the exact location of the diffusivity disruption at millimetre level.
\end{abstract}

### 2018-04-25 contents/ap_paper.tex
Updates from ShareLaTeX

\graphicspath{{thesis/images/ap-reg/}}

\chapter{Study II: Diffusion Weighted Image to T1 Co-registration Without Reverse-blip: Investigation of Best Practices}
\chaptermark{Study II}
\label{section:study2}

@@

\begin{figure}[p]
\includegraphics[width=\linewidth]{figure4}
\caption{Visual comparisons of MDWI to T1 co-registrations. 
Shown are examples of MDWI images from one subject deformed using different registration transforms derived from the different image intermediates. 
T1 is overlaid as thresholded outlines in red. 
Primary differences in registration quality can be found in the brainstem (left and middle columns), and in the lateral ventricles, especially in the ventricular-caudate boundary (right column). }
\centering
\label{fig:APfigure4}
\end{figure}

@@

    \begin{figure}[p]
    \includegraphics[width=\linewidth]{figure5}
    \caption{Similarity scores between deformed dMRI scalar images and T1 following registrations with CC as cost function for SyN. 
    Three steps of the registration process are plotted: Initial identity transformation (none), affine transformation (affine) and affine+SyN transformation (SyN).
    Top panel shows similarity scores as measured by CC and MI, and bottom panel shows average percent change progressions of CC and MI across dMRI scalar images from \protect $T_{\text{Affine}}$ to \protect $T_{\text{Identity}}$, and from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Affine}}$. The error bars denote standard deviation across subjects.}
    \centering
    \label{fig:APfigure5}
    \end{figure}

@@

        \begin{figure}[p]
        \includegraphics[width=\linewidth]{figure6}
        \caption{
        Percent changes of similarity score from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Identity}}$ with CC as cost function. Top row shows the violin plot distribution of CC and MI percent change under All Deformations and Autodeformations. Bottom row plots the statistically significant pair-wise correlations of the different transforms. 
        Under All Deformations (\protect $D_{\text{all}}$), CC percent change of FA is shown to be significantly lower (p<0.05) than that of all other transform types. GFA can be observed to have higher percent change than that of B0, MDWI, and FA. Under Autodeformations (\protect $D_{\text{auto}}$), CC percent change of FA and GFA are significantly higher than others, while AP has lower percent change. MI shows little percentage between all the images; GFA shows significantly differences under \protect $D_{\text{auto}}$, but also shows high variability.}
        \centering
        \label{fig:APfigure6}
        \end{figure}
    \end{landscape}

@@

    \begin{figure}[p]
    \includegraphics[width=\linewidth]{sup1}
    \caption{Similarity scores between deformed dMRI scalar images and T1 following registrations with MI as cost function for SyN. 
    Three steps of the registration process are plotted: Initial identity transformation (none), affine transformation (affine) and affine+SyN transformation (SyN).
    Top panel shows similarity scores as measured by CC and MI, and bottom panel shows average percent change progressions of CC and MI across dMRI scalar images from \protect $T_{\text{Affine}}$ to \protect $T_{\text{Identity}}$, and from \protect $T_{\text{Affine}}$+SyN to\protect  $T_{\text{Affine}}$.}
    \centering
    \label{fig:APSup1}
    \end{figure}

@@

    \thispagestyle{empty}% empty page style (?)
    \begin{landscape}% Landscape page
        \begin{figure}[p]
        \includegraphics[width=\linewidth]{sup2}
        \caption{Percent changes of similarity score from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Identity}}$ with MI as cost function. Top row shows the violin plot distribution of CC and MI percent change under All Deformations and Autodeformations. Bottom row plots the statistically significant pair-wise correlations of the different transforms.
        Under all deformations ( \protect$D_{\text{all}}$), CC percent change of FA is shown to be significantly lower (p<0.05) from that of other transform types. In autodeformations (\protect$D_{\text{auto}}$). FA has higher percent change. MI percent changes show little differences, and once again highlights the high variability of GFA. }
        \centering
        \label{fig:APSup2}
        \end{figure}
    \end{landscape}
}

>>>>

\graphicspath{{images/ap-reg/}}

\chapter{Study II: Diffusion Weighted Image Co-registration: Investigation of Best Practices}
\chaptermark{Study II}
\label{section:study2}

@@

\begin{figure}[p]
\includegraphics[width=\linewidth]{figure4}
\caption[Visual comparisons of MDWI to T1 co-registrations.]{Visual comparisons of MDWI to T1 co-registrations. 
Shown are examples of MDWI images from one subject deformed using different registration transforms derived from the different image intermediates. 
T1 is overlaid as thresholded outlines in red. 
Primary differences in registration quality can be found in the brainstem (left and middle columns), and in the lateral ventricles, especially in the ventricular-caudate boundary (right column). }
\centering
\label{fig:APfigure4}
\end{figure}

@@

    \begin{figure}[p]
    \includegraphics[width=\linewidth]{figure5}
    \caption[Similarity scores between deformed dMRI scalar images and T1 following registrations with CC as cost function for SyN.]{Similarity scores between deformed dMRI scalar images and T1 following registrations with CC as cost function for SyN. 
    Three steps of the registration process are plotted: Initial identity transformation (none), affine transformation (affine) and affine+SyN transformation (SyN).
    Top panel shows similarity scores as measured by CC and MI, and bottom panel shows average percent change progressions of CC and MI across dMRI scalar images from \protect $T_{\text{Affine}}$ to \protect $T_{\text{Identity}}$, and from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Affine}}$. The error bars denote standard deviation across subjects.}
    \centering
    \label{fig:APfigure5}
    \end{figure}

@@

        \begin{figure}[p]
        \includegraphics[width=\linewidth]{figure6}
        \caption[Percent changes of similarity score from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Identity}}$ with CC as cost function.]{
        Percent changes of similarity score from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Identity}}$ with CC as cost function. Top row shows the violin plot distribution of CC and MI percent change under All Deformations and Autodeformations. Bottom row plots the statistically significant pair-wise correlations of the different transforms. 
        Under All Deformations (\protect $D_{\text{all}}$), CC percent change of FA is shown to be significantly lower (p<0.05) than that of all other transform types. GFA can be observed to have higher percent change than that of B0, MDWI, and FA. Under Autodeformations (\protect $D_{\text{auto}}$), CC percent change of FA and GFA are significantly higher than others, while AP has lower percent change. MI shows little percentage between all the images; GFA shows significantly differences under \protect $D_{\text{auto}}$, but also shows high variability.}
        \centering
        \label{fig:APfigure6}
        \end{figure}
    \end{landscape}

@@

    \begin{figure}[p]
    \includegraphics[width=\linewidth]{sup1}
    \caption[Similarity scores between deformed dMRI scalar images and T1 following registrations with MI as cost function for SyN. ]{Similarity scores between deformed dMRI scalar images and T1 following registrations with MI as cost function for SyN. 
    Three steps of the registration process are plotted: Initial identity transformation (none), affine transformation (affine) and affine+SyN transformation (SyN).
    Top panel shows similarity scores as measured by CC and MI, and bottom panel shows average percent change progressions of CC and MI across dMRI scalar images from \protect $T_{\text{Affine}}$ to \protect $T_{\text{Identity}}$, and from \protect $T_{\text{Affine}}$+SyN to\protect  $T_{\text{Affine}}$.}
    \centering
    \label{fig:APSup1}
    \end{figure}

@@

    \thispagestyle{empty}% empty page style (?)
    \begin{landscape}% Landscape page
        \begin{figure}[p]
        \includegraphics[width=\linewidth]{sup2}
        \caption[Percent changes of similarity score from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Identity}}$ with MI as cost function.]{Percent changes of similarity score from \protect $T_{\text{Affine}}$+SyN to \protect $T_{\text{Identity}}$ with MI as cost function. Top row shows the violin plot distribution of CC and MI percent change under All Deformations and Autodeformations. Bottom row plots the statistically significant pair-wise correlations of the different transforms.
        Under all deformations ( \protect$D_{\text{all}}$), CC percent change of FA is shown to be significantly lower (p<0.05) from that of other transform types. In autodeformations (\protect$D_{\text{auto}}$). FA has higher percent change. MI percent changes show little differences, and once again highlights the high variability of GFA. }
        \centering
        \label{fig:APSup2}
        \end{figure}
    \end{landscape}
}

### 2018-04-25 contents/paper1.tex
Updates from ShareLaTeX

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure1.png}
\caption{Figure shows the MR images of a typical MS-TN patient at the level of pons. The presence of MS plaque in the CNS at the level of trigeminal nerve is characteristic of MS-TN (panel A). Panel B shows the expected course of the trigeminal nerve (CN V) as it synapses onto the principal trigeminal and motor trigeminal nuclei. Note that the exact anatomical relationship between the MS plaque and CN V in this instance cannot be clearly discerned in T2 due to the lack of contrast between the nerve and the surrounding tissue (panels A,B,C). Panels C and D show the placements of the ROIs along the CN V nerve bilaterally in order to measure diffusivity statistics, overlaid with the axial anatomical/ DTI image. The colors of the DTI image are rendered as color-by-orientation, where by convention red represents left–right, greens represents anterior–posterior and blue represents inferior–superior orientations. Diffusion statistics were measured from four groups of regions: cisternal segment, root entry zone (REZ), pontine, and peri-lesional, as illustrated. Analogous regions in TN patients and controls were similarly placed. Care was taken to differentiate pontine and peri- lesional ROIs in order to measure lesioned and unlesioned regions.}
\centering
\label{fig:MSfigure1}
\end{figure}

@@

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure2.png}
\caption{Reconstruction of the trigeminal nerves at the level of pons from a single healthy subject in superior axial views. The reconstructed tracts are overlaid onto the axial DTI/anatomical images. The colors of the underlying scan are rendered as color-by-orientation, where by convention red represents left–right, greens represents anterior–posterior and blue represents inferior–superior orientations. The colors of the tractography fibers represent the spectrum values of FA (0 to 1), as shown by the legend at the top right corner. Panel A shows that SDT tractography delineated the cerebellar peduncle fibers, but cannot distinguish the brainstem trigeminal fibers (yellow arrows). Red arrows denote the starting point of tract generation for both SDT and MTT methods. Panel B shows that MTT tractography visualized CN V as it coursed through the brainstem towards the trigeminal main sensory nuclei.}
\centering
\label{fig:MSfigure2}
\end{figure}

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure3.png}
\caption{Comparison of tractography delineation between affected and unaffected CN V nerve in the region of the trigeminal main sensory nucleus using the MTT method. Images are shown in superior axial view. Panel A shows the MS-TN tractography delineations. There is focal decrease in CN V FA in the region of the MS lesion (yellow arrow). Panel B shows the TN tractography delineations, where there is no such FA decrease in the brainstem segment. Please refer to Figure \protect\ref{fig:MSfigure2} for explanations of color scales and figure annotations.}
\centering
\label{fig:MSfigure3}
\end{figure}

@@

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure4.png}
\caption{Top row: Diffusivity difference estimate matrices of the Cisternal ROIs demonstrating significant differences in regional diffusivities between MS-TN patients, TN patients and controls. Colored squares indicate statistically significant differences (p\textless 0.05) after correcting for multiple comparisons with false discovery rate. The (+/-) sign indicate symptomatic (+) and asymptomatic (-) sides of the subject group. The matrix should be read as ‘Y axis in relation to X axis’—for example, A red square indicates that the vertical axis measure (e.g. TN+, the affected side in TN patients) is significantly higher than that of the horizontal axis measure (e.g. MS-TN+, the affected side in MS-TN patients); blue indicates that the measure is lower. The number within the square indicates relative estimated differences. Bottom row: Box-plots of the diffusivity distributions. Note that the control group does not have pain lateralities, therefore ipsilateral indicate the right side in the control group in this figure. The mean of the left-right diffusivities in controls were used for comparisons against other groups. The analysis highlights TN+ specific diffusivity differences in the Cisternal segment of the trigeminal nerve. Specifically TN+ has significantly higher FA than all the other groups, and that there are significant differences between TN+ and TN-. The points above and below the box plots are outliers. Significances are not indicated in the plot, for pair-wise significance, please refer to the estimate matrices above row.}
\centering
\label{fig:MSfigure4}
\end{figure}

>>>>

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure1.png}
\caption[Figure shows the MR images of a typical MS-TN patient at the level of pons.]{Figure shows the MR images of a typical MS-TN patient at the level of pons. The presence of MS plaque in the CNS at the level of trigeminal nerve is characteristic of MS-TN (panel A). Panel B shows the expected course of the trigeminal nerve (CN V) as it synapses onto the principal trigeminal and motor trigeminal nuclei. Note that the exact anatomical relationship between the MS plaque and CN V in this instance cannot be clearly discerned in T2 due to the lack of contrast between the nerve and the surrounding tissue (panels A,B,C). Panels C and D show the placements of the ROIs along the CN V nerve bilaterally in order to measure diffusivity statistics, overlaid with the axial anatomical/ DTI image. The colors of the DTI image are rendered as color-by-orientation, where by convention red represents left–right, greens represents anterior–posterior and blue represents inferior–superior orientations. Diffusion statistics were measured from four groups of regions: cisternal segment, root entry zone (REZ), pontine, and peri-lesional, as illustrated. Analogous regions in TN patients and controls were similarly placed. Care was taken to differentiate pontine and peri- lesional ROIs in order to measure lesioned and unlesioned regions.}
\centering
\label{fig:MSfigure1}
\end{figure}

@@

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure2.png}
\caption[Reconstruction of the trigeminal nerves at the level of pons from a single healthy subject in superior axial views.]{Reconstruction of the trigeminal nerves at the level of pons from a single healthy subject in superior axial views. The reconstructed tracts are overlaid onto the axial DTI/anatomical images. The colors of the underlying scan are rendered as color-by-orientation, where by convention red represents left–right, greens represents anterior–posterior and blue represents inferior–superior orientations. The colors of the tractography fibers represent the spectrum values of FA (0 to 1), as shown by the legend at the top right corner. Panel A shows that SDT tractography delineated the cerebellar peduncle fibers, but cannot distinguish the brainstem trigeminal fibers (yellow arrows). Red arrows denote the starting point of tract generation for both SDT and MTT methods. Panel B shows that MTT tractography visualized CN V as it coursed through the brainstem towards the trigeminal main sensory nuclei.}
\centering
\label{fig:MSfigure2}
\end{figure}

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure3.png}
\caption[Comparison of tractography delineation between affected and unaffected CN V nerve in the region of the trigeminal main sensory nucleus using the MTT method.]{Comparison of tractography delineation between affected and unaffected CN V nerve in the region of the trigeminal main sensory nucleus using the MTT method. Images are shown in superior axial view. Panel A shows the MS-TN tractography delineations. There is focal decrease in CN V FA in the region of the MS lesion (yellow arrow). Panel B shows the TN tractography delineations, where there is no such FA decrease in the brainstem segment. Please refer to Figure \protect\ref{fig:MSfigure2} for explanations of color scales and figure annotations.}
\centering
\label{fig:MSfigure3}
\end{figure}

@@

\begin{figure}[p]
\includegraphics[width=\textwidth]{figure4.png}
\caption[Diffusivity difference estimate matrices of the Cisternal ROIs demonstrating significant differences in regional diffusivities between MS-TN patients, TN patients and controls.]{Top row: Diffusivity difference estimate matrices of the Cisternal ROIs demonstrating significant differences in regional diffusivities between MS-TN patients, TN patients and controls. Colored squares indicate statistically significant differences (p\textless 0.05) after correcting for multiple comparisons with false discovery rate. The (+/-) sign indicate symptomatic (+) and asymptomatic (-) sides of the subject group. The matrix should be read as ‘Y axis in relation to X axis’—for example, A red square indicates that the vertical axis measure (e.g. TN+, the affected side in TN patients) is significantly higher than that of the horizontal axis measure (e.g. MS-TN+, the affected side in MS-TN patients); blue indicates that the measure is lower. The number within the square indicates relative estimated differences. Bottom row: Box-plots of the diffusivity distributions. Note that the control group does not have pain lateralities, therefore ipsilateral indicate the right side in the control group in this figure. The mean of the left-right diffusivities in controls were used for comparisons against other groups. The analysis highlights TN+ specific diffusivity differences in the Cisternal segment of the trigeminal nerve. Specifically TN+ has significantly higher FA than all the other groups, and that there are significant differences between TN+ and TN-. The points above and below the box plots are outliers. Significances are not indicated in the plot, for pair-wise significance, please refer to the estimate matrices above row.}
\centering
\label{fig:MSfigure4}
\end{figure}

### 2018-04-25 contents/reviews/pain.tex
Updates from ShareLaTeX

\textit{Nociceptors}, the peripheral nociceptors respond to noxious stimuli. These receptors have free nerve endings, and can respond to thermal, mechanical, and/or chemical stimuli, such as physical trauma or heat/cold temperatures. These receptors broadly separate into two categories, differentiated by the fibre type. Acute nociceptors are associated with $A\delta$ fibres, which permit fast signal conductance (5 to 30 m/s) and trigger a prickly or sharp pain sensation; examples include needle prick other acute trauma. Dull pain is triggered by receptors that are associated with slow conducting C fibres ($<1$ m/s). $A\delta$ and C fibres can be co-activated in the presence of noxious stimuli. C-fibre nociceptors can be further differentiated into mechano-heat, silent, and polymodal subtypes. $A\delta$-fibre nociceptors can be divided into high-threshold mechano-heat, low-threshold mechano-heat, and noxious cold subtypes.

@@

The peripheral nociception signals are conducted from the axonal endings of the nociception neurons, whose cell bodies are located in the spinal cord dorsal root gangions, to the dorsal horn of the spinal cord (\ref{fig:drg}). The $A\delta$ and C afferents project directly to lamina I of the dorsal horn. $A\delta$ afferents also directly project to lamina V. Lamina II, or the substantia gelatinosa, also receives inputs from C afferents, which indirectly synapse to lamina I neurons through interneurons. The dorsal horn also contains a dense layer of local interneurons in lamina III and IV, and receive input from innocuous somatosensory afferents at these layers.

@@

In each spinal afferent level, the dorsal horn neurons project noceciptive signals to the brain via five distinct paths (Figure \ref{fig:pain-pathways}). Lamina I, V-VII axons crosses the mid-line to form the \textit{spinothalamic} tract that ascend in the contralateral anterolateral column of the spinal cord, and project to thalamic nuclei. Laminae VII and VIII neurons from the \textit{spinorecticular} tract ascend via the anterolateral quadrant of the spinal cord, without crossing the midline, to reach the reticular formation and thalamus. The \textit{spinomesencephalic} tract include axons from laminae I and V, and ascend in the anterolateral quadrant to the mesencephalic reticular formation, including the periaqueductual gray. The \textit{cervicothalamic} tract include only the white matters of the upper two cervical segments. It takes inputs from laminae III and IV, and projects to the thalamus via the medial lemniscus. The spinohypothalamic tract contains axons from laminae I, V, and VIII, and project to hypothalamic nuclei for autonomic functions.

@@

Peripheral neuropathic pains can arise when there are disruptions in peripheral signal transmission. The disruption can be a result of injury to the first order peripheral afferent neurons, which result in increased sensitivity to stimuli and via ectopic action potentials. Changes in peripheral neurotropic factors as a result of an injury can cause increased membrane excitability or spurious sprouting of the synaptic terminals in the dorsal horn of the spinal cord. Additionally, maladaptive changes may occur in the secondary sensory neurons and interneurons in the dorsal-horn laminae that amplifies or reroutes the noxious signals \cite{Elmes2004}.

>>>>

\textit{Nociceptors}, the peripheral nociceptors respond to noxious stimuli. These receptors have free nerve endings, and can respond to thermal, mechanical, and/or chemical stimuli, such as physical trauma or heat/cold temperatures. These receptors broadly separate into two categories, differentiated by the fibre type. Acute nociceptors are associated with $A\delta$ fibres, which permit fast signal conductance (5 to 30 m/s) and trigger a prickly or sharp pain sensation; examples include needle prick other acute trauma. Dull pain is triggered by receptors that are associated with slow conducting C fibres ($<1$ m/s). C-fibre nociceptors can be further differentiated into mechano-heat, silent, and polymodal subtypes. $A\delta$-fibre nociceptors can be divided into high-threshold mechano-heat, low-threshold mechano-heat, and noxious cold subtypes \cite{Julius2013}.

@@

The peripheral nociception signals are conducted from the axonal endings of the nociceptive neurons, whose cell bodies are located in the spinal cord dorsal root gangions, to the dorsal horn of the spinal cord (\ref{fig:drg}). The $A\delta$ and C afferents project directly to lamina I of the dorsal horn. $A\delta$ afferents also directly project to lamina V. Lamina II, or the substantia gelatinosa, also receives inputs from C afferents, which indirectly synapse to lamina I neurons through interneurons. The dorsal horn also contains a dense layer of local interneurons in lamina III and IV, and receive input from innocuous somatosensory afferents at these layers.

@@

In each spinal afferent level, the dorsal horn neurons project noceciptive signals to the brain via five distinct paths (Figure \ref{fig:pain-pathways}). Lamina I, V-VII axons crosses the mid-line to form the \textit{spinothalamic} tract that ascend in the contralateral anterolateral column of the spinal cord, and project to thalamic nuclei. Laminae VII and VIII neurons from the \textit{spinorecticular} tract ascend via the anterolateral quadrant of the spinal cord, without crossing the midline, to reach the reticular formation and thalamus. The \textit{spinomesencephalic} tract include axons from laminae I and V, and ascend in the anterolateral quadrant to the mesencephalic reticular formation, including the periaqueductual gray. The \textit{cervicothalamic} tract include only the white matters of the upper two cervical segments. It takes inputs from laminae III and IV, and projects to the thalamus via the medial lemniscus. The spinohypothalamic tract contains axons from laminae I, V, and VIII, and project to hypothalamic nuclei for autonomic functions \cite{kandel2000principles}.

@@

Peripheral neuropathic pains can arise when there are disruptions in peripheral signal transmission. The disruption can be a result of injury to the first order peripheral afferent neurons, which result in increased sensitivity to stimuli and via ectopic action potentials \cite{Costigan2009}. Changes in peripheral neurotropic factors as a result of an injury can cause increased membrane excitability or spurious sprouting of the synaptic terminals in the dorsal horn of the spinal cord. Additionally, maladaptive changes may occur in the secondary sensory neurons and interneurons in the dorsal-horn laminae that amplifies or reroutes the noxious signals \cite{Elmes2004}.

### 2018-04-25 contents/sagit.tex
- reorg folder
Updates from ShareLaTeX

\graphicspath{{images/sagit/drafts/}}

@@

\begin{figure}[p]
\centering
\includegraphics[width=140mm]{figure1.png}
\caption{Flow diagram of the SAGIT group tractography framework. The SAGIT framework is designed to be fully configurable and extensible.}
\label{fig:SGfigure1}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[width=\linewidth]{figure2.png}
\caption{Example of the auto-generated visual panels. The panels that best represent the resulting anatomy are compiled. For anatomical reference, the best representative anatomy image slice is composed with the tract conjunction map. The color scale of the conjunction image represents the percentage reproducible for each region, where 1=100\%. For each algorithm (rows), different visual perspectives are created. Together 6 different anatomies were delineated (columns). To see the full visual reports generated, please see supplemental.}
\label{fig:SGfigure2}
\end{figure}

@@

\afterpage{%
    \clearpage% Flush earlier floats (otherwise order might not be correct)
    \thispagestyle{empty}% empty page style (?)
    \begin{landscape}% Landscape page
\begin{figure}[p]
\includegraphics[width=\linewidth]{figure3.png}
\caption{Example of merged tractography models for each of the techniques and anatomy. The fornix is colored to distinguish tracts from each of the 42 subjects, in order to demonstrate the effect of combining the group data. The other anatomies are coloured by FA intensity. The specific FA measures from each point were sampled from their native DWI space before tractography deformation. The grouping of the similar FA measures across the subjects show that the registrations were accurate. The Red Nuclei shows different metrics (FA and AD) depending on lateralization, in order to highlight the pontine decussation.}
\centering
\label{fig:SGfigure3}
\end{figure}
    \end{landscape}
    \clearpage% Flush page
}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{figure5.png}
	\caption{Normalized anatomical scores as rated by experts. The error bars denote standard deviations of the ratings. The fornix serves as the control anatomy, therefore the represented filtered and unfiltered image are identical. It can be observed that unfiltered anatomy increases rater variability in some and decrease in others.}
	\label{fig:SGfigure5}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figure6.png}
\caption{NOS scores for each anatomy. It can be noted that having ROI filters severely reduces reproducibility in deterministic methods.}
\label{fig:SGfigure6}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figure7.png}
\caption{Correlation between rating scores and NOS. Filtered tractography NOS is significantly (p\textless 0.05) correlated with ratings. However unfiltered tractography show no significance.}
\label{fig:SGfigure7}
\end{figure}

>>>>

\graphicspath{{images/sagit/}}

@@

\begin{figure}[p]
\centering
\includegraphics[width=140mm]{figure1.pdf}
\caption{Flow diagram of the SAGIT group tractography framework. The SAGIT framework is designed to be fully configurable and extensible.}
\label{fig:SGfigure1}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[width=\linewidth]{figure2.png}
\caption[Example of the auto-generated visual panels.]{Example of the auto-generated visual panels. The panels that best represent the resulting anatomy are compiled. For anatomical reference, the best representative anatomy image slice is composed with the tract conjunction map. The color scale of the conjunction image represents the percentage reproducible for each region, where 1=100\%. For each algorithm (rows), different visual perspectives are created. Together 6 different anatomies were delineated (columns). To see the full visual reports generated, please see supplemental.}
\label{fig:SGfigure2}
\end{figure}

@@

\afterpage{%
    \clearpage% Flush earlier floats (otherwise order might not be correct)
    \thispagestyle{empty}% empty page style (?)
    \begin{landscape}% Landscape page
\begin{figure}[p]
\includegraphics[width=\linewidth]{figure3.png}
\caption[Example of merged tractography models for each of the techniques and anatomy.]{Example of merged tractography models for each of the techniques and anatomy. The fornix is colored to distinguish tracts from each of the 42 subjects, in order to demonstrate the effect of combining the group data. The other anatomies are coloured by FA intensity. The specific FA measures from each point were sampled from their native DWI space before tractography deformation. The grouping of the similar FA measures across the subjects show that the registrations were accurate. The Red Nuclei shows different metrics (FA and AD) depending on lateralization, in order to highlight the pontine decussation.}
\centering
\label{fig:SGfigure3}
\end{figure}
    \end{landscape}
    \clearpage% Flush page
}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{figure5.pdf}
	\caption[Normalized anatomical scores as rated by experts.]{Normalized anatomical scores as rated by experts. The error bars denote standard deviations of the ratings. The fornix serves as the control anatomy, therefore the represented filtered and unfiltered image are identical. It can be observed that unfiltered anatomy increases rater variability in some and decrease in others.}
	\label{fig:SGfigure5}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figure6.pdf}
\caption{NOS scores for each anatomy. It can be noted that having ROI filters severely reduces reproducibility in deterministic methods.}
\label{fig:SGfigure6}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figure7.pdf}
\caption{Correlation between rating scores and NOS. Filtered tractography NOS is significantly (p\textless 0.05) correlated with ratings. However unfiltered tractography show no significance.}
\label{fig:SGfigure7}
\end{figure}
